\newpage
\section{Module 07 - {\ttfamily ""module07\_sponsor"}}
\label{sec:module07_predicate}

\subsection{Overview}
Module 07 handles gas sponsorship operations from the owners ZapWallet. It validates and processes EIP-712 signed
gas sponsorship parameters, allowing gas UTXOs to be securely used for asset exchange, free usage, or sponsorship signatures
to be cancellation.\\


The module operates in three modes:
\begin{itemize}
\item \textbf{Sponsor}: Trade gas for another asset within configurable tolerance bounds
\item \textbf{Gaspass}: Allow free gas usage with guaranteed return amounts
\item \textbf{Cancel}: Move a gas UTXO without requiring exchange
\end{itemize}

Formally, for a transaction $T$ with gas sponsorship operation $O$, the validation is:\\

$[ v: O \rightarrow {true, false} \text{ where } O = (w, p) ]$ \\

where $v$ is the validation function, $O$ is the sponsorship operation tuple , $w$ is the EIP-712 signature
witness and $p$ are the sponsorship parameters\\


This allows gas UTXOs to be used flexibly and securely from the owners ZapWallet, enabling use cases like
transaction sponsorship, free trials, and cancellations while enforcing the sponsor's signed intent.\\

\subsection{Sponsorship Parameters}
The gas sponsorship parameters are defined by the \texttt{GasSponsor} struct:
\begin{lstlisting}
    struct GasSponsor {
    command: String,
    returnaddress: b256,
    inputgasutxoid: b256,
    expectedgasoutputamount: u256,
    expectedoutputasset: b256,
    expectedoutputamount: u256,
    tolerance: u256,
}
\end{lstlisting}

Where:
\begin{itemize}
\item \texttt{command}: Operation type (\texttt{"sponsor"}, \texttt{"gaspass"}, or \texttt{"cancel"})
\item \texttt{returnaddress}: Expected gas return address
\item \texttt{inputgasutxoid}: UTXO ID of the input gas coin
\item \texttt{expectedgasoutputamount}: Expected gas return amount
\item \texttt{expectedoutputasset}: Expected asset ID for exchange (\texttt{"sponsor"} only)
\item \texttt{expectedoutputamount}: Expected asset amount for exchange (\texttt{"sponsor"} only)
\item \texttt{tolerance}: Acceptable deviation from expected exchange amount (\texttt{"sponsor"} only)
\end{itemize}

These parameters are hashed according to EIP-712 and signed by the gas owner, providing secure authorization for the
specified usage of their gas UTXO.\\


\subsection{Validation Flow}
The core validation logic follows these steps:

\begin{enumerate}
\item Extract EIP-712 signature witness
\item Collect and categorize transaction inputs and outputs
\item Validate the specified gas UTXO is present and owned by signer
\item Process outputs according to \texttt{command}:
\begin{itemize}
\item \texttt{"sponsor"}: Verify asset exchange amounts are within tolerance
\item \texttt{"gaspass"}: Verify gas return amount
\item \texttt{"cancel"}: Validate gas UTXO movement
\end{itemize}
\item Rebuild \texttt{GasSponsor} struct and verify EIP-712 signature
\end{enumerate}

If all checks pass, the transaction is approved.\\



\subsection{Asset Flow}
A transaction containing Module 07 processes the following assets:

\begin{itemize}
\item \textbf{Gas UTXO}: The input gas coin (\text{BASE\_ASSET}) owned by the owners ZapWallet specified by \texttt{inputgasutxoid}
\item \textbf{Exchange Asset} (\texttt{"sponsor"} only): The asset traded to the gas owner
\end{itemize}

For the three different types of transactions \texttt{"sponsor"}, \texttt{"gaspass"} and \texttt{"cancel"} the asset flow is described below.\\


% % ----------------------------------------------------------------------------------------------
% sponsor
\subsubsection{\texttt{"sponsor"} Asset Flow}

The gas input UTXO must be present and owned by the owner of the ZapWallet. For \texttt{"sponsor"}
operations, the output exchange asset must match the \texttt{expectedoutputasset} and \\
\texttt{expectedoutputamount} and no \texttt{tolerance} is specified. Any remaining gas not explicitly output is returned
to the owner via a Change output.\\

\textbf{Transaction Structure}:
\begin{figure}[H]
    \scalebox{0.85}{
    \centerline{
    \xymatrix@R=2.5pc@C=1.5pc{
        & \txt{Input type\\ Asset type\\ Amount}
        & \txt{Verification\\ Signature \\ type}
        & \txt{Output type\\ Receiver\\ Asset type\\ Amount} \\
        & *+[F]\txt{Coin predicate:\\ Module 07\\ asset: Module 07\\ AssetId\\ amount: 1}
            \ar[r] & *+[H]\txt{Gas Sponsor\\ Validation}
            \ar[r] & *+[F=]\txt{type: Coin\\ to: Module 07\\ asset: Module 07\\ AssetId\\ amount: 1} \\
        & *+[F:purple]\txt{Coin:  User\\ asset: Other Coin AssetId\\ amount: $n$,\\ where $n \geq 1$}
            \ar[r] & *+[H]\txt{VM ownership validation \\ \& EIP712 Validation}
            \ar[r] & *+[F]\txt{type: Coin\\ to: Master\\ asset: Other Coin AssetId\\ amount: $\text{expectedoutputamount}$} \\
        & \phantom{*+[F=]\txt{Placeholder}}
            & \phantom{*+[F=]\txt{Placeholder}}
            \ar[r] & *+[F:purple]\txt{type: Change\\ to: User\\ asset: Other Coin AssetId\\ amount: $n - \text{expectedoutputamount}$} \\
        % ----------------------
        & *+[F=]\txt{Coin predicate: Master\\ asset: BASE\_ASSET\\ amount: value}
            \ar[r] & *+[H]\txt{Master Validation \&\\ EIP712 Validation}
            \ar[r] & *+[F=]\txt{type: Coin\\ to: Master\\ asset: BASE\_ASSET\\ amount: value - gas cost} \\
        & \phantom{*+[F=]\txt{Placeholder}}
            & \phantom{*+[F=]\txt{Placeholder}}
            \ar[r] & *+[F=]\txt{type: Change\\ to: Master\\ asset: BASE\_ASSET\\ amount: NA } \\
        % & \phantom{*+[F=]\txt{Placeholder}}
        %     & \phantom{*+[F=]\txt{Placeholder}}
        %     & *+[F.:grey:dashed]\txt{type: Change\\ to: master\\ asset: BASE\_ASSET\\ amount: \\ $\text{receiver value} - \text{network fee} - \text{network fee}$}
    }}
    }
    \caption{Module 07 Gas Sponsorship Transaction Flow for \texttt{"sponsor"}}
    \label{fig:module07-flow}
\end{figure}



% % ----------------------------------------------------------------------------------------------
% gasspass
\subsubsection{\texttt{"gasspass"} Asset Flow}

The gas input UTXO must be present and owned by the owner of the ZapWallet. For \texttt{"gasspass"}
operations, no exchange asset is required by the third party using the owners ZapWallet gas UTXO. The
\texttt{expectedgasoutputamount} gas return amount is specified by the owner. No \texttt{tolerance} is required. Any remaining gas not explicitly output is returned
to the owner via a Change output.\\




\textbf{Transaction Structure}:
\begin{figure}[H]
    \scalebox{0.85}{
    \centerline{
    \xymatrix@R=2.5pc@C=1.5pc{
        & \txt{Input type\\ Asset type\\ Amount}
        & \txt{Verification\\ Signature \\ type}
        & \txt{Output type\\ Receiver\\ Asset type\\ Amount} \\
        & *+[F]\txt{Coin predicate:\\ Module 07\\ asset: Module 07\\ AssetId\\ amount: 1}
            \ar[r] & *+[H]\txt{Gas Sponsor\\ Validation}
            \ar[r] & *+[F]\txt{type: Coin\\ to: Module 07\\ asset: Module 07\\ AssetId\\ amount: 1} \\
        % ----------------------
        & *+[F=]\txt{Coin predicate:\\ Master\\ asset: BASE\_ASSET\\ amount:\\ value}
            \ar[r] & *+[H]\txt{Master Validation \&\\ EIP712 Validation}
            \ar[r] & *+[F=]\txt{type: Coin\\ to: Master\\ asset: BASE\_ASSET\\ amount: $\text{expectedgasoutputamount}$} \\
        & \phantom{*+[F=]\txt{Placeholder}}
            & \phantom{*+[F=]\txt{Placeholder}}
            \ar[r] & *+[F=]\txt{type: Change\\ to: Master\\ asset: BASE\_ASSET\\ amount: NA } \\
        % & \phantom{*+[F=]\txt{Placeholder}}
        %     & \phantom{*+[F=]\txt{Placeholder}}
        %     & *+[F.:grey:dashed]\txt{type: Change\\ to: master\\ asset: BASE\_ASSET\\ amount: \\ $\text{receiver value} - \text{network fee} - \text{network fee}$}
    }}
    }
    \caption{Module 07 Gas Sponsorship Transaction Flow for \texttt{"gasspass"}}
    \label{fig:module07-flow}
\end{figure}


% % ----------------------------------------------------------------------------------------------
% cancel
\subsubsection{\texttt{"cancel"} Asset Flow}

For \texttt{"cancel"} operations, the UTXO of the owners gas is simply spent and sent back to the owner Zapwallet. This renders an already
signed \texttt{"sponsor"} or \texttt{"gasspass"} signature invalid. \\


\textbf{Transaction Structure}:
\begin{figure}[H]
    \scalebox{0.85}{
    \centerline{
    \xymatrix@R=2.5pc@C=1.5pc{
        & \txt{Input type\\ Asset type\\ Amount}
        & \txt{Verification\\ Signature \\ type}
        & \txt{Output type\\ Receiver\\ Asset type\\ Amount} \\
        & *+[F]\txt{Coin predicate:\\ Module 07\\ asset: Module 07\\ AssetId\\ amount: 1}
            \ar[r] & *+[H]\txt{Gas Sponsor\\ Validation}
            \ar[r] & *+[F]\txt{type: Coin\\ to: Module 07\\ asset: Module 07\\ AssetId\\ amount: 1} \\
        % ----------------------
        & *+[F=]\txt{Coin predicate:\\ Master\\ asset: BASE\_ASSET\\ amount:\\ value}
            \ar[r] & *+[H]\txt{Master Validation \&\\ EIP712 Validation}
            \ar[r] & *+[F=]\txt{type: Change\\ to: Master\\ asset: BASE\_ASSET\\ amount: NA} \\
    }}
    }
    \caption{Module 07 Gas Sponsorship Transaction Flow for \texttt{"cancel"}}
    \label{fig:module07-flow}
\end{figure}





% % ----------------------------------------------------------------------------------------------
%

\subsection{Security Properties}
\subsubsection{Signature Integrity}
The EIP-712 signature is verified against the rebuilt \texttt{GasSponsor} struct, ensuring that all
validated parameters match the signer's intent exactly. The signing domain and struct type hash are
fixed at deployment, preventing any modification.\\

$[ s = \text{Sign}(H(GasSponsor), d_{owner}) ]$
$[ \text{Verify}(s, H(GasSponsor), d_{contract}) = v ]$\\

Where:
\begin{itemize}
\item $s$: The EIP-712 signature
\item $H$: The EIP-712 struct hash function
\item $d_{owner}$: The gas owner's signing domain
\item $d_{contract}$: The contract's verification domain
\item $v$: The boolean validation result
\end{itemize}


\subsubsection{Asset Conservation}


For command $c$, gas amount $g$, and exchange amount $e$:
\texttt{command = "sponsor"}:\\

$[ \sum \text{outputs} = (g - r) + (e \pm t) + r ]$

\texttt{command = "gaspass"}:\\

$[ \sum \text{outputs} = (g - r) + r ]$

\texttt{command = "cancel"}:\\

$[ \sum \text{outputs} = g ]$

This ensures that assets are never lost or created, only moved according to the
validated sponsorship parameters.





\subsubsection{Asset Conservation}
The total output amounts are always equal to the input amounts, ensuring no assets are lost or created.
For a gas sponsorship transaction with gas amount $g$, expected return $r$, exchange amount $e$, and tolerance $t$:

\textbf{Sponsor Operation:}
$[ \sum \text{outputs} = \underbrace{(g - r)}{\text{Unused gas}} + \underbrace{(e \pm t)}{\text{Exchange asset}} + \underbrace{r}_{\text{Returned gas}} = g + (e \pm t) ]$

\textbf{Gaspass Operation:}
$[ \sum \text{outputs} = \underbrace{(g - r)}{\text{Unused gas}} + \underbrace{r}{\text{Returned gas}} = g ]$

\textbf{Cancel Operation:}
$[ \sum \text{outputs} = \underbrace{g}_{\text{Moved gas}} ]$

In each case, the output amounts are conserved:
\begin{itemize}
\item \textbf{Sponsor}: The sum of unused gas, exchanged asset, and returned gas equals the input gas plus the exchange asset amount (within tolerance).
\item \textbf{Gaspass}: The sum of unused and returned gas equals the input gas amount.
\item \textbf{Cancel}: The output gas amount is equal to the input gas amount.
\end{itemize}

This asset conservation ensures that the gas sponsorship operations only move assets according to the validated parameters, without any loss or unauthorized creation of funds.






















\subsubsection{Replay Prevention}
The \texttt{inputgasutxoid} parameter uniquely identifies the exact gas UTXO authorized for the operation.
This ID is signed along with the other parameters, preventing any modification or replay of the sponsorship.


\subsection{Error Handling}
The module includes error handling that propagates errors to at maximum the \text{main()} scope. However as predicates
must return boolean results, all errors result in a \text{false} returned from the main execution.



