\newpage
\section{Module 00 - {\ttfamily "module00\_upgrade"}}
\label{sec:module00_predicate}

\subsection{Overview}
Module 00 serves as the upgrade predicate for a ZapWallet, it facilitating secure transitions between Zap wallet versions. This
module ensures that upgrades are authorized by the wallet owner and maintains proper asset flow during the upgrade process. The
transaction structure for the upgrade procedure can be gas sponsored by a third party or gas can be spent by the owner of the ZapWallet.

\subsection{Module 00 Core Functionality}
The module validates upgrade transactions through:
\begin{itemize}
    \item Verification of upgrade acknowledgment signatures
    \item Asset flow validation
    \item Version control checks
    \item Ownership verification
\end{itemize}


\subsection{Configuration Parameters}
The module requires the following configuration:
\begin{itemize}
    \item \texttt{OWNER\_ADDRESS}: Wallet owner's EVM address
    \item \texttt{NONCE\_ASSETID}: Wallet's nonce AssetId
    \item \texttt{MODULE\_KEY00\_ASSETID}: Upgrade module's AssetId
    \item \texttt{ZAPMANAGER\_V1}: ZapManager V1 contract Address
    \item \texttt{VERSION}: Module00 version identifier
\end{itemize}



\subsection{Asset Requirements}
An upgrade transaction must satisfy the following asset conditions:
\begin{itemize}
    \item Module 00 asset must be sent to the ZapManager V1 contract
    \item Nonce assets must be present in both inputs and outputs and be amount accounted for
    \item Nonce asset ownership must be verified in the transaction flow
\end{itemize}

\subsection{Upgrade Acknowledgment}
The upgrade process requires the owner to sign a structured acknowledgment message that contains crucial upgrade parameters:

\begin{lstlisting}
    WalletUpgradeAcknowledgment {
        from_address: b256,     // Current wallet master address
        to_address: b256,       // Target wallet master address
        current_version: String, // Fixed at "1.0.0" for V1
        target_version: String,  // Target version in X.Y.Z format
        utxo_id: b256          // UTXO ID of the upgrade module
    }
\end{lstlisting}

These parameters are formatted into a human-readable message that the owner must sign:

\begin{lstlisting}
    "UPGRADE ACKNOWLEDGMENT:
    I understand and authorize that:\n\
    1. ALL assets (Tokens, NFTs, and ETH) will be transferred from 0x<from_address> to 0x<to_address>
    2. This upgrade will change my wallet from version <current_version> to <target_version>
    3. This action cannot be reversed once executed
    4. This authorization is valid only for this specific upgrade request
    5. One time upgrade UTXO ID: <utxo_id>"
\end{lstlisting}

The explicit acknowledgment ensures the owner understands the scope and implications of the upgrade process.



\subsection{Validation Flow}
The module performs the following verification steps:
\begin{enumerate}
    \item Verifies module00 asset is properly sent to ZapManager V1
    \item Confirms nonce asset presence and ownership in inputs
    \item Validates nonce asset destination in outputs
    \item Constructs and verifies the upgrade acknowledgment message
    \item Recovers signer address from EIP-191 signature
    \item Validates signer against wallet owner address
\end{enumerate}

\subsection{Security Properties}
The module enforces several critical security properties:




\subsubsection{Asset Conservation}
The upgrade module must ensure proper handling of the module asset during the upgrade process. For the module00
asset $M_0$ and ZapManager contract address $Z$, the following must be true:
\[ \exists \text{ output } o : o.asset = M_0 \land o.to = Z \]
This means there must exist a transaction output that sends the module00 asset to the ZapManager contract
address. This requirement ensures the module asset is properly transferred for the upgrade process and cannot be diverted elsewhere.

\subsubsection{Nonce Continuity}
The nonce asset $N$ must maintain proper ownership throughout the
upgrade. For addresses $(from, to)$:
\[ \exists \text{ input } i : i.asset = N \land i.owner = from \]
\[ \exists \text{ output } o : o.asset = N \land o.to = to \]
This enforces two key requirements:
\begin{itemize}
   \item The transaction must include the nonce asset from the owner's current wallet address
   \item The nonce asset must be properly forwarded and the amount be equal.
\end{itemize}
This mechanism prevents transaction replay attacks in Modules 01-03 and ensures proper ZapWallet upgrade.

\subsubsection{Signature Verification}
The upgrade must be authorized by the wallet owner through a cryptographic signature. For owner
address $O$ and upgrade acknowledgment message $m$:
\[ ec\_recover(signature, hash(m)) = O \]
This validation:
\begin{itemize}
   \item Recovers the signer's address from the EIP-191 signature
   \item Verifies it matches the wallet owner's address
   \item Ensures only the legitimate owner can authorize upgrades
\end{itemize}
The message $m$ contains critical upgrade parameters including source and destination addresses,
ensuring the owner explicitly acknowledges the upgrade details.




\subsection{Upgrade Transaction Inputs and Output}
The following diagram illustrates the transaction inputs, verification mechanism for each transaction input and resulting outputs. If
the upgrade transaction is sponsored then the total amount of ZapWallet owner \text{BASE\_ASSET} input should be equal to the \text{BASE\_ASSET}
output going to the V2 master. The Sponsor should pay for the gas to the transaction. Otherwise if the user is spending their own gas execute
the upgrade transaction the \text{BASE\_ASSET} input should equal the \text{BASE\_ASSET} output minus the network fee.


\begin{figure}[H]
    \scalebox{0.85}{
    \centerline{
    \xymatrix@R=2.5pc@C=1.5pc{
        & \txt{Input type\\ Asset type\\ Amount}
        & \txt{Verification\\ Signature\\ type}
        & \txt{Output type\\ Receiver\\ Asset type\\ Amount} \\
        & *+[F]\txt{Coin predicate:\\ Module 00\\ asset: Module 00\\ AssetId\\ amount: $u$\\ where $u \geq 1$}
            \ar[r] & *+[H]\txt{Upgrade\\ acknowledgement\\ EIP-191}
            \ar[r] & *+[F]\txt{type: Coin\\ to: ZapManager\\ contract\\ asset: Module 00\\ AssetId\\ amount: $u$} \\
        & *+[F=]\txt{Coin predicate:\\ Master\\ asset: Nonce AssetId\\ amount: $n$,\\ where $n \geq 1$}
            \ar[r] & *+[H]\txt{Asset\\ Ownership\\ Validation}
            \ar[r] & *+[F=]\txt{type: Coin\\ to: V2 Master\\ asset: Nonce AssetId\\ amount: $n - 1$} \\
        & *+[F=]\txt{Coin predicate:\\ Master\\ asset: BASE\_ASSET\\ amount: x}
            \ar[r] & *+[H]\txt{VM Balance \\ Validation }
            \ar[r] & *+[F=]\txt{type: Coin\\ to: V2 Master\\ asset: BASE\_ASSET\\ amount: remaining \\ or x if not sponsored} \\
        & *+[F][orange]\txt{Coin: Sponsor\\ asset: BASE\_ASSET\\ amount: $sg$}
            \ar[r] & *+[H][orange]\txt{VM Balance\\ Validation}
            \ar[r] & *+[F][orange]\txt{type: Change\\ to: Sponsor\\ asset: BASE\_ASSET\\ amount:\\ $sg$ remaining}
    }}
    }
    \caption{Module 00 Upgrade Transaction Flow and Verification}
    \label{fig:upgrade-flow-xy}
\end{figure}




